Parsing Postfix log files

### 2007-10-16 doc/logparser.tex
Improved the cached/discarded section, referring to the normal/shuffle/reverse section.
Moved the rule hits graph to a subsection of its own, and added three more
graphs giving an improved breakdown of the rule hits, as in the overall graph
it's difficult to distinguish what's happening in the lower end.

Extended the introduction and conclusion with the new material.

Graphs are an excellent means of displaying data, transforming a
meaningless stream of numbers into an easily comprehensible form, where
anomalies and patterns are immediately obvious.  These graphs are used to
illustrate the topics discussed in section~\ref{rule efficiency}.  The
graphs in the first section cover parser scalability, demonstrating that
performance scales linearly with input size.  The impact of rule ordering
is shown in section~\ref{rule ordering graphs}, and the lower impact of
rule ordering on larger log files is explained.  The final group of graphs
vividly shows the huge impact that caching compiled regexs has on parser
performance.

@@

The Y axis represents the following:

@@

@@

Two large dips can be seen in graph~\ref{normal regex vs disarded regex
factor} at log files 22 and 62--68, corresponding to the spikes in log file
size in graph~\ref{execution time vs file size vs number lines}.  The
anomalies are caused by a combination of a mail loop and a delivery failure
which resulted in a huge number of log lines, all of which were matched by
a single rule.  Sorting the rules by the number of successful matches has
the effect that this rule will be tried first, so the average number of
regexs compiled per line drops sharply for those eight log files, resulting
in a much lower increase in relative execution time, as detailed below.

@@

@@

@@

The graphs presented in this section illustrate the topics discussed in
section~\ref{rule efficiency}.  The first collection of graphs are about
parser scalability, showing the linearly relationship between execution
time and input size.  Section~\ref{rule ordering graphs} relates to the
effect of rule ordering on execution time, and the unexpected consequences
of specific inputs.  The necessity of caching compiled regexs is attested
to by the final group of graphs, where the difference between caching and
discarding is staggering.

>>>>

Graphs are an excellent means of displaying data, transforming a
meaningless stream of numbers into an easily comprehensible form, where
anomalies and patterns are immediately obvious.  These graphs are used to
illustrate the topics discussed in section~\ref{rule efficiency}.  The
graphs in the first section cover parser scalability, demonstrating that
performance scales linearly with input size.  The impact of rule ordering
is shown in section~\ref{rule ordering graphs}, and the anomalous dips and
peaks apparent in some graphs are explained.  The next group of graphs
vividly shows the huge impact that caching compiled regexs has on parser
performance.  This section concludes with a breakdown of the rule hits
accumulated during a single test run of the parser.

@@

The Y axis in the graph above represents the following:

@@

The following graphs show the impact that not caching compiled regexs has
on parser performance: on typical log files the execution time when not
caching compiled regexs is 500--600\% of the execution time when caching;
reversing the perspective shows that cached execution time is merely
17--20\% of non-cached execution time.  Caching compiled regexs is probably
the single most effective optimisation possible in the parser's
implementation.

@@

Two large dips can be seen in graph~\ref{normal regex vs disarded regex
factor} at log files 22 and 62--68, corresponding to the spikes in log file
size in graph~\ref{execution time vs file size vs number lines}.  The
reason for the anomalous log files has already been explained in
section~\ref{Why are there dips in the graphs?}.

The distribution of log lines across rules is much different and the
average number of rules consulted per log line is much lower when there is
a mail loop; this results in far fewer regex compilations per line than
when there isn't a mail loop, and a correspondingly decreased execution
time.  The increases in execution time when not caching regexs for
different combinations of log files are summarised in the table below:

@@

The number of hits per rule is quite unevenly spread, resembling a Power
Law distribution~\cite{powerlaw}.

@@

As the graph above is quite difficult to read it has been separated into
three sections below: low, middle and high.

\noindent\includegraphics{build/plot-hits-low.ps}

It is apparent from the low graph that some rules have few or no hits;
those with zero hits are rules which were written to parse log files used
during development of the parser but not utilised in the test runs
performed for this document.

\noindent\includegraphics{build/plot-hits-middle.ps}

\noindent\includegraphics{build/plot-hits-high.ps}

@@

The graphs presented in this section illustrate the topics discussed in
section~\ref{rule efficiency}.  The first collection of graphs are about
parser scalability, showing the linearly relationship between execution
time and input size.  Section~\ref{rule ordering graphs} relates to the
effect of rule ordering on execution time, and the unexpected consequences
of specific inputs.  The necessity of caching compiled regexs is attested
to by the third group of graphs, where the difference between caching and
discarding is staggering.  This section concluded with a breakdown of the
rule hits accumulated during a single test run of the parser.
